---
title:"Reproducible Research: Peer Assessment 1"
output:
html_document:
  keep_md: true
---

## Loading and preprocessing the data

```{r}
library(lattice)
#These commands read the data into R
setwd("C:/Users/juane_000/Desktop")
x<-read.csv("activity.csv")
names(x)
```

## What is mean total number of steps taken per day?

```{r}
y<-tapply(x$steps,as.factor(x$date),sum)
hist(y,xlab="Steps per day",breaks=10,main="Number of steps per day",col="blue")
t<-mean(y,na.rm=TRUE)
u<-median(y,na.rm=TRUE)
```

The mean is `r t` and the median is `r u`

## What is the average daily activity pattern?

```{r}
z<-tapply(na.rm=TRUE,x$steps,as.factor(x$interval),mean)
plot(names(z),z,type="l",xlab="Interval Index",ylab="Average number of steps",main="Average number of steps 
per interval  of  time",col="green")
v<-names(z[z==max(z)])[1]
```

The maximum average daily activity occurs on the interval `r v`, which has the maximum average number of steps.
In other words, the maximum average daily activity occurs from 8:35 AM to 9:00 AM.

## Imputing missing values 

```{r}
a<-length(x$steps[is.na(x$steps)])
```

There are `r a` missing values in this dataset.

The following algorithm will modify the dataset by replacing the missing values( of a missing day) with the median of `r u` (obtained previously) divided by the number of intervals in that day (288 intervals) so that the sum of the missing values of the respective missing day sum up to `r u`.

```{r}
     for(i in 1:length(as.factor(levels(x$date)))){
       w<-x[x$date==as.factor(levels(x$date))[i],]
         if(length(w[,1][!is.na(w[,1])])==0){
           for(j in 1:length(w[,1])){
             w[j,1]<-u/length(w[,1])}}else{
             for(j in 1:length(w[,1])){
               w[j,1]<-x[x$date==as.factor(levels(x$date))[i],][j,1]}}
               x[x$date==as.factor(levels(x$date))[i],]<-w}
```

Now, the histogram looks like this:

```{r}
c<-tapply(x$steps,as.factor(x$date),sum)
hist(c,xlab="Steps per day",breaks=10,main="Number of steps per day",col="red")
d<-mean(c,na.rm=TRUE)
e<-median(c,na.rm=TRUE)
```

The histogram looks similar to the one obtained with missing values. The resulting mean is `r d` and the resulting median is `r e`. These results are equal to the previous results obtained with missing values.

## Are there differences in activity patterns between weekdays and weekends?

Let's add another variable to our list which specifies if a given day is weekday or weekend.

```{r}
 q<-rep(0,length(x$date))
     x<-cbind(x,q)
     for(i in 1:length(as.factor(levels(x$date)))){
       w<-x[x$date==as.factor(levels(x$date))[i],]
      for(j in 1:length(w[,1])){
       r<-weekdays(as.POSIXlt(w$date))
       if(r[j]=="Saturday" |r[j]=="Sunday"){w[j,4]<-"Weekend"}else{w[j,4]<-"Weekday"} 
      }
       x[x$date==as.factor(levels(x$date))[i],]<-w}
```

Now, let's compute the mean steps per interval for weekdays and weekend days:

```{r}
f<-x[x$q!="Weekday",]
g<-x[x$q!="Weekend",]
m<-tapply(f$steps,f$interval,mean)
n<-tapply(g$steps,g$interval,mean)
par(mfrow=c(1,2))
plot(names(m),m,type="l",xlab="Interval",ylab="Average Steps per Interval",main="Average steps  per Interval 
for Weekends",col="cyan")
plot(names(n),n,type="l",xlab="Interval",ylab="Average Steps per Interval",main="Average steps  per Interval
for Weekdays",col="orange")
```